---
title: "EGPIT Update, July 2 2026"
format:
  html:
    embed-resources: true
editor: visual
---

**Note: this markdown had become obsolete bc the package functions have been updated from the versions used prior to the meeting. Forgot to commit before some of those changes. HTML is the version presented. Recommend not rendering this file.**

Comments from the May 7 meeting, James email, or the developers to do list are provided as heading below. Each heading is followed by an example for the figure modification.

```{r, message = FALSE, warning = FALSE, include = FALSE}
library(EGprocess)
library(tidyverse)
library(flextable)
```

### Make EG lines discontinuous (Done)

### Add figure borders (Done)

```{r, fig.width = 6, fig.height= 4}
#| fig.format: png
#| fig.dpi: 300
brood_Igushik <- get_brood(run_data = data_Igushik)

plot_escapement(brood_data = brood_Igushik, 
                goal_data = goal_Igushik, 
                title = "Igushik River Sockeye Salmon")
```

### Scale plot axes (Example developed)

```{r, fig.width = 6, fig.height= 4}
#| fig.format: png
#| fig.dpi: 300

source(".\\plot_escapement_scaled.R")
plot_escapement(brood_data = brood_Igushik, 
                goal_data = goal_Igushik, 
                title = "Igushik River Sockeye Salmon")
```

```{r, fig.width = 6, fig.height= 4}
#| fig.format: png
#| fig.dpi: 300
brood_lt10K <- brood_Igushik %>% mutate(across(S:R, function(x) x/1e3))
goal_lt10K <- goal_Igushik %>% mutate(across(lb:ub, function(x) x/1e3))
plot_escapement(brood_lt10K,
                goal_lt10K,
                "max escapement is below 10K"
)
```

```{r, fig.width = 6, fig.height= 4}
#| fig.format: png
#| fig.dpi: 300
brood_gt10K<- brood_Igushik %>% mutate(across(S:R, function(x) x/1e2))
goal_gt10K<- goal_Igushik %>% mutate(across(lb:ub, function(x) x/1e2))
plot_escapement(brood_gt10K,
                goal_gt10K,
                "max escapement is above 10K"
)
```

### Figure caption improvements (Done)

### Add EG range to figure captions (Done)

```{r, fig.width = 6, fig.height= 5}
#| fig.format: png
#| fig.dpi: 300
post_Igushik <- c(post_Igushik_byr63_05, post_Igushik_byr63_15)

plot_ey(post_Igushik,
        brood_Igushik,
        goal_dat = goal_Igushik,
        "Igushik River Sockeye Salmon",
        multiplier = 1e-5)
```

```{r, fig.width = 6, fig.height= 5}
#| fig.format: png
#| fig.dpi: 300
goal_imaginary <-
  goal_Igushik %>%
  bind_rows(
    data.frame(yr = 2026,
               lb = 175000,
               ub = 500000)
  )
goal_imaginary

plot_ey(post_Igushik,
        brood_Igushik,
        goal_dat = goal_imaginary,
        "Imaginary River Sockeye Salmon",
        new_finding = TRUE,
        multiplier = 1e-5)
```

### Comment: Table CV's and referecne point rounding (Done)

```{r}
#| fig.format: png
#| fig.dpi: 300
table_SR(post_Igushik[2], title = "Igushik River Sockeye Salmon", multiplier = 1e-6)
```

### Comment: visual legends for presentations (example developed)

```{r, fig.width = 6, fig.height= 6.5, fig.cap = "Relationship between escapement and recruitment and the spawning escapement that maximizes sustained yield for Igushik River Sockeye salmon."}
#| fig.format: png
#| fig.dpi: 300
source(".\\plot_spawner_recruit_legend.R")
# No goal change, single analysis
plot_SR(post_Igushik[2],
        brood_Igushik,
        goal_dat = goal_Igushik,
        "Igushik River Sockeye Salmon",
        multiplier = 1e-5)
```

```{r, fig.width = 6, fig.height= 6.5, warning = FALSE}
#| fig.format: png
#| fig.dpi: 300
# No goal change, two analyses
plot_SR(post_Igushik,
        brood_Igushik,
        goal_dat = goal_Igushik,
        "Igushik River Sockeye Salmon",
        multiplier = 1e-5)
```

```{r, fig.width = 6, fig.height= 6.5, warning = FALSE}
#| fig.format: png
#| fig.dpi: 300
#new finding
plot_SR(post_Igushik,
        brood_Igushik,
        goal_dat = goal_imaginary,
        "Imaginary River Sockeye Salmon",
        new_finding = TRUE,
        multiplier = 1e-5)
```

Plan to move towards a combination of visual and text legends for each figure. Also need to migrate all figures to Times New Roman.
