---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EGprocess

<!-- badges: start -->
<!-- badges: end -->

The R package EGprocess facilitates simple and standard creation of escapement goal data reporting. 

## Package Installation

You can install the development version of EGprocess from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("adamreimer/EGprocess")
```

## Overview
EGprocess provides a standardized set of R functions to be used by ADF&G staff during 
the escapement goal review process. This package is designed to help staff from 
both fisheries divisions and all regions produce consistent, high-quality figures 
and tables to be used as part of the  Board of Fisheries (BOF) meetings.
Salmon escapement goals in Alaska are reviewed every three years, aligned with 
the BOF’s regulatory meeting cycle. EGprocess supports this work by creating 
standardized outputs for spawner-recruit analyses, posterior summaries, 
expected yield curves, optimal yield profiles (OYP), and comparisons among 
updated and historical escapement goal findings.
EGprocess can be used directly in R, or as a downstream companion to the 
Pacific Salmon SR Escapement Goal Analysis Shiny Application ("Hamachan Shiny App"), 
which produces the input data structures used by the package.

### Why Standardize Figures?
Standardizing figure structure, captions, and themes reduces the cognitive load 
on reviewers, allowing them to focus on interpretation instead of formatting 
differences across regions or divisions. EGprocess ensures:
 - Consistent plot titles (Stock Name + Species Name).
 - Standardized layout, theming, and color schema.
 - Concise, reproducible figure notes that act as built-in legends.
 - Reproducible workflows for future users and archiving. 

Standardization does not prevent customization. Staff can treat EGprocess outputs 
as templates, modifying elements as needed for unique stock-specific situations. 
In more complex cases, users can build new plots from scratch using the tools provided by EGprocess.

### Package Structure / Workflow
EGprocess starts with MCMC output data created either by the Hamachan Shiny App, 
or during personalized spawner-recruit analyses using rJAGS. Specifically, 
the data must contain nodes of lnalpha, beta, phi, and sigma.  
There are repeated elements within EGprocess for ease of use. Common data inputs include:  
 - *brood_data*: Brood table created by function `make_brood()`  
 - *goal_data*: Current and historical escapement goals  
 - *posterior_data*: MCMC samples for ln(alpha), beta, sigma, and related parameters  
 - *profile_data*: Data output from function `get_profile()`   


### Flexible MSY Targets
While the typical escapement goal benchmark is 90% of MSY, it may be necessary 
to demonstrate other MSY levels. For example, the included dataset for 
Igushik River sockeye salmon shows a a near-zero probability of achieving 
90% MSY at the lower bound of its existing goal and historical management 
reasoning for a suboptimal but stable yield.  
In cases like this, EGprocess supports side-by-side evaluation of optimal and 
suboptimal yields. Users can specify a suboptimal target directly in the 
`get_profile()` function and generate profile plots that evaluate escapement goals
relative to both targets.

### Updating Escapement Goals 
Most frequently, researchers are tasked with updating an existing escapement goal
with additional returns. In such cases, it is helpful to compare data since the 
previous assessment to the updated dataset. EGprocess facilitates simple comparison 
between such datasets so that researchers can assess the validity of the goal in 
light of additional data. The plotting function `plot_profile_facet()` shows
the Optimal Yield Profile from both datasets.  


## Function Overview

### get_profile()
Function `get_profile()` generates optimal yield profiles (OYP) and 
expected yield curves from posterior MCMC samples.
```{r} 
library(EGprocess)
data("post_Igushik_byr63_15")
get_profile(posterior_data = post_Igushik_byr63_15, multiplier = 1e-5)
``` 


### make_age()
Function `make_age()` creates age-composition proportions (A3, A4, etc) from 
run data for use in brood‑table construction.
```{r}
make_age(age_data = data_Igushik, min_age = 3, max_age = 8) 
```

### make_brood()
Function `make_brood()` builds the brood table by combining escapement, 
recruitment, and age-composition information.
```{r}
p_Igushik <- make_age(data_Igushik, min_age = 3, max_age = 8)
make_brood(data = data_Igushik, p = p_Igushik)
```

### plot_profile()
Function `plot_profile()` generates a standardized Optimal Yield Profile (OYP) 
plot from a single profile dataset.
```{r}
Igushik_profile <- get_profile(post_Igushik_byr63_15, multiplier = 1e-5)
plot_profile(profile_data = Igushik_profile, goal_data = goal_Igushik, 
title = "Igushik River Sockeye Salmon")
```

### 
Function `plot_profile_facet()` produces faceted OYP plots for comparing 
multiple posterior datasets (e.g., updated vs. historical). 
```{r}
post_list <- list(
     'Brood Years: 1963-2005' = post_Igushik_byr63_05,
     'Brood Years: 1963-2015' = post_Igushik_byr63_15)
profile_list <- lapply(post_list, get_profile, multiplier = 1e-5)

plot_profile_facet(profile_data = profile_list, goal_data = goal_Igushik,
title = "Igushik River Sockeye Salmon")
```
### 
Function 
```{r}

```
### 
Function 
```{r}

```
### 
Function 
```{r}

```



## Included Datasets


