---
title: "EGprocess: Standard Process"
output:
  html_document:   # for HTML with figure numbering 
    number_sections: false
    fig_caption: true
    self_contained: true
vignette: >
  %\VignetteIndexEntry{Standard}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "../man/figures/Standard-",
  out.width = "100%"
)
```

```{r, message = FALSE, warning = FALSE, include = FALSE}
library(EGprocess)
library(tidyverse)
library(flextable)
```

# Overview

In addition to standardizing common tables and figures between regions and divisions, `EGprocess` also attempts to standardize information and visual reference points between figures that are typically presented together. Some commonalities you will find in all figures produced by this package:

-   Use of a table and figure titles. Titles include the stock and species being displayed in the following format: Stock Name, Species Name.
-   Subtitles or labels specifying the brood years included in the analysis. Brood year information is conveyed through items names in `posterior_list` or `profile_list` and follow the format: Brood years: xxx-yyy
-   Consistent use of visual clues across plots.
    -   Analysis: dashed lines represent the original analysis (i.e. the analysis used to establish the current escapement goal) while solid lines represent the updated analysis. 
    -   Escapement goals: Grey fill represents the current escapement goal range and, when applicable, brown fill represents new escapement goal range.
    -   Empirical data: Hollow circles represent data included in the original analysis and filled circles represent data added to the updated analysis.
    -   Visual reference: Dotted lines.
-   Textual legends: Notes associated with each figure act so the key to each plot's visual cues. This approach maximizes the percentage of the figure dedicated to displaying the data, ensures the legend travels with the figure, and allows for brief figure captions.
-   Use of Arial font.

The following 2 sections will demonstrate these unifying design principles across 2 common scenarios following an escapement goal analysis update. 

# Finding: No Change

Often escapement goal analyses are updated but the updated analysis continues to support the current escapement goal range. This section will demonstrate the production of tables and figures in that scenario. 

## Historical Escapement Plot {.tabset}

EGPIT decided historical escapements should be depicted as a bar plot overlain with the contemporary escapement goal range and color coded to indicate escapements below the lower bound of the contemporary escapement goal. *Note: The EGPIT representation is visually pleasing. An alternative is offered which is admittedly less aesthetic but has the advantage of matching the visual cues used in the other plots.*

### EGPIT
```{r historicS_nochange, fig.width = 6, fig.height= 4}
brood_Igushik <- get_brood(run_data = data_Igushik)

plot_escapement(brood_data = brood_Igushik, 
                goal_data = goal_Igushik, 
                title = "Igushik River Sockeye Salmon")
```

### Alternative
```{r, echo = FALSE}
plot_alt_escapement <- function(brood_data,
                             goal_data,
                             title,
                             new_finding = FALSE,
                             broods_updated = 0){
  brood_data <- brood_data %>% dplyr::filter(!is.na(S))
  
  yr_max <- if(max(goal_data$yr) < max(brood_data$yr)){max(brood_data$yr + 2)}else{max(goal_data$yr) + 2}
  goal <-
    goal_data %>%
    dplyr::mutate(yr = yr - 0.5,
                  dplyr::across(c(lb, ub), function(x){ifelse(is.na(x), -99, x)}),
                  finding = ifelse(yr == max(goal_data$yr) - 0.5 & isTRUE(new_finding), "new", "current")) %>%
    tidyr::complete(yr = tidyr::full_seq(c(yr, yr_max), 0.1)) %>%
    tidyr::fill(lb, ub, finding, .direction = "down") %>%
    dplyr::mutate(across(c(lb, ub), function(x){ifelse(x == -99, NA, x)}))
  
  cap <- 
  case_when(isTRUE(new_finding) & broods_updated >= min(goal_data$yr) ~ stringr::str_wrap("Note: Filled 
    circles indicated observations added to the dataset since the escapement goal last changed. Escapement 
    goal ranges are shaded, brown for the new escapement goal finding and grey for current and historical goals.", 
    width = 85),
    isTRUE(new_finding) & broods_updated < min(goal_data$yr)  ~ stringr::str_wrap("Note: Escapement 
    goal ranges are shaded, brown for the new escapement goal finding and grey for current and historical goals.", 
    width = 85),
    !isTRUE(new_finding) & broods_updated >= min(goal_data$yr) ~ stringr::str_wrap("Note: Filled 
    circles indicated observations added to the dataset since the escapement goal last changed. Escapement 
    goal ranges are shaded grey", width = 85),
    !isTRUE(new_finding) & broods_updated < min(goal_data$yr) ~ stringr::str_wrap("Note: Escapement goal ranges 
    are shaded grey.", width = 85)
  )
                   
  
  brood_data %>%
    dplyr::select(yr, S) %>%
    dplyr::mutate(update = ifelse(yr >= broods_updated, "updated", "existing")) %>%
    ggplot2::ggplot(aes(x = yr)) +
    ggplot2::geom_point(aes(y = S, shape = update)) +
    ggplot2::geom_line(aes(y = S), linewidth = 0.5) +
    ggplot2::geom_ribbon(ggplot2::aes(x = yr, ymin = lb, ymax = ub, fill = finding),
                         data = goal,
                         alpha = 0.5,
                         inherit.aes = FALSE) +
    ggplot2::scale_fill_manual(values = c("new" = "#AB7E4C", "current" = "gray80")) +
    ggplot2::scale_shape_manual(values = c("updated" = 16, "existing" = 1)) +
    ggplot2::scale_y_continuous(labels = scales::comma) +
    ggplot2::labs(title = title,
                  x = "Year",
                  y = "Escapement",
                  caption = cap) +
    theme_eg()
  
}
```
```{r alternateS_nochange, fig.width = 6, fig.height= 4}
plot_alt_escapement(brood_Igushik,
                    goal_Igushik,
                    "Igushik River Sockeye Salmon",
                    broods_updated = 2014)
```

## Spawner-Recruit Plot{-}

EGPIT decided spawner-recruit (SR) relationships should depicted the SR pairs used to estimate the relationship (with color coding to indicate SR pairs included in the original analysis and added to the updated analysis), the estimated SR relationship, $S_{MSY}$, and the current goal range.

```{r SRplot_nochange, fig.width = 6, fig.height= 6}
post_Igushik <- c(post_Igushik_byr63_05, post_Igushik_byr63_15)

plot_SR(posterior_list = post_Igushik, 
        brood_data = brood_Igushik, 
        goal_data = goal_Igushik, 
        title = "Igushik River Sockeye Salmon", 
        multiplier = 1e-5)
```

## Expected Yield Plot
EGPIT decided expected yield plots should depicted the empirical data used to estimate the relationship (with color coding to indicate SR pairs added to the dataset since the last time the escapement goal was changed), the estimate of expected yield, and the current goal range. 

```{r ey_nochange, fig.width = 6, fig.height= 5}
plot_ey(posterior_list = post_Igushik, 
        brood_data = brood_Igushik, 
        goal_data = goal_Igushik, 
        title = "Igushik River Sockeye Salmon", 
        multiplier = 1e-5)
```

## Optimal Yield Profile

EGPIT decided escapement goals intended to target MSY should use a universal standard target yield of greater than 90% of MSY. When producing OYP plots this means a single line, overlain with the current escapement goal is needed to describe the probability of maximizing yield within the escapement goal range. By displaying the OYP generated from both the original and updated analyses ADF&G staff can easily compare how the empirical data added since the last escapement goal change has impacted the relationship between MSY and the goal. In this case, 10 additional brood years resulted in a slight rightward shift of the OYP and minimal changes to the probability of achieving at least 90% of MSY at both the upper and lower bound of the existing goal. Thus, this presentation supports the escapement goal committee's recommendation of no change to the escapement goal. 

```{r OYP_nochange, fig.width = 6, fig.height = 6}
profile_Igushik <- get_profile(post_Igushik, multiplier = 1e-5)

plot_profile(profile_Igushik,
             goal_Igushik,
             "Igushik River Sockeye Salmon")
```

## Spawner-Recruit Parameter Table

EGPIT determined the following parameters should be presented after each escapement goal update. Notice that the function only expects the posterior from the updated analysis. *note to EGPIT: I'd say it's a statistical best practice to present confidence intervals instead of SEs but I don't recall EGPIT having an opinion on this point.*

```{r SRtable_nochange}
table_SR(post_Igushik[2], 
         title = "Igushik River Sockeye Salmon", 
         multiplier = 1e-6)
```

# Finding: New Escapement Goal Range

During the escapement goal review process for Igushik River Sockeye salmon the escapement goal review committee elected not to change the escapement goal. For illustrative purposes we will imagine they had recommended an escapement goal change and examine the changes to `EGprocess` outputs in that scenario. 

Imagine the committee had disregarded the stability of the relationship between the escapement goal bounds and the OYP probabilities displayed in Figure 5 and decided the goal was too aggressive relative to a target of 90% of MSY. This imaginary committee may have decided increases to both the lower and upper bounds of the existing escapement goal were appropriate. This decision could have been communicated using `EGprocess` outputs by adding an additional row to the `goal_data` file containing the new escapement goal finding and the year the new finding would go into effect.

In addition to the modified `goal_data` argument, the argument `new_finding` is set equal to TRUE for the following plots.  In what follows we will also change the figure titles to make clear that this is an imaginary example. Table 1 is unchanged and will be omitted from this section. 

```{r}
goal_imaginary <-
  goal_Igushik %>%
  bind_rows(
    data.frame(yr = 2026,
               lb = 175000,
               ub = 500000)
  )
goal_imaginary
```

## Historical Escapement Plot {.tabset}

Following the EGPIT standard the new escapement goal finding is shown as a change to the escapement goal range which extends into the future. *Note: The alternative is similar but highlights the presence of a new finding using color coding.*

### EGPIT
```{r historicS_change, fig.width = 6, fig.height= 4}
plot_escapement(brood_data = brood_Igushik, 
                goal_data = goal_imaginary, 
                title = "Igushik River Sockeye Salmon")
```

### Alternative
```{r, alternateS_change, fig.width = 6, fig.height= 4}
plot_alt_escapement(brood_Igushik,
                    goal_imaginary,
                    "Imaginary River Sockeye Salmon",
                    broods_updated = 2014, 
                    new_finding = TRUE)
```

## Spawner-Recruit and Expected Yield Plots {-}

When a new escapement goal finding is made the spawner-recruit plot is identical except the new finding is displayed on the plot and highlighted using color coding. Color coding is extremely helpful here as the new finding can be difficult to detect unless the reader remembers the current and new escapement goal bounds very accurately. The expected yield plot (not shown) is similarly modified.

```{r SRplot_change, fig.width = 6, fig.height= 6}
plot_SR(posterior_list = post_Igushik, 
        brood_data = brood_Igushik, 
        goal_data = goal_imaginary, 
        title = "Imaginary River Sockeye Salmon",
        new_finding = TRUE,
        multiplier = 1e-5)
```

## Optimal Yield Profile

When a new escapement goal finding is made the new finding is overlain on the optimal yield profile for the updated analysis while the current escapement goal range is overlain on the optimal yield profile from the original analysis. This configuration allows staff to compare the degree of similarity between the new and current escapement goals, relative to the analysis responsible for each finding, with respect to maximizing MSY. In this case, the new escapement goal finding offers improved symmetry between the upper on lower escapement goal bounds and roughly similar probability of achieving at least 90% of MSY at the lower bound but decreased probability of achieving at least 90% of MSY at the upper bound. 

```{r OYP_change, fig.width = 6, fig.height = 6}
profile_Igushik <- get_profile(post_Igushik, multiplier = 1e-5)

plot_profile(profile_Igushik,
             goal_imaginary,
             "Imaginary River Sockeye Salmon",
             new_finding = TRUE)
```
