---
title: "EGprocess: Standard Process"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: FALSE
    number_sections: false
    fig_caption: true
    self_contained: true
vignette: >
  %\VignetteIndexEntry{Standard}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "../man/figures/Standard-",
  out.width = "100%"
)
```

```{r, message = FALSE, warning = FALSE, include = FALSE}
library(EGprocess)
library(tidyverse)
library(flextable)
```

This vignette is the second in a series describing use of the `EGprocess` package and covers common situations which arise when updating an escapement goal. New users may want to refer to the vignette about [package structure](EGprocess-Orientation.html). Advanced users may want to refer to the vignette about [special cases](EGprocess-Special_cases.html).

# Overview

In addition to standardizing common tables and figures between regions and divisions, `EGprocess` also attempts to standardize information and visual reference points between figures that are typically presented together. Some commonalities you will find in all figures produced by this package:

-   Table and figure titles in the following format: Stock Name, Species Name
-   Arial font
-   Figure borders
-   Current escapement goal ranges and/or new escapement goal findings shown in figures are also specified as numbers 
-   Number scaling: When the maximum number being plotted is between 10,000 and 1,000,000, axis labels are scaled to display thousands (e.g. 2K). The scale changes to millions (e.g. 2M) when the maximum number being plotted is greater than 1,000,000.

In addition, figures associated with spawner-recruit analyses are further standardized as follows:

-   Indication of the brood years included in the analysis
-   Consistent visual clues
    -   Analysis: dashed lines represent the original analysis while solid lines represent the updated analysis
    -   Escapement goals: Grey fill represents the current escapement goal range, brown fill represents the new escapement goal range
    -   Empirical data: Hollow circles represent data included in the original analysis and filled circles represent data added to the updated analysis
    -   Reference lines are dotted
-   Self contained legends[^1] 

[^1]: Traditionally, information not contained in the figure legend was included in the figure caption associated with the figure. In `EGprocess`, information not clarified by the figure legend is included as text within the figure border. This approach ensures all of the needed information 'travels' together with the figure.

The following sections will demonstrate these unifying design principles across 2 common scenarios following an escapement goal analysis update. 

# Finding: No Change

During the 2026 escapement goal review process for Igushik River Sockeye salmon the escapement goal review committee elected not to change the escapement goal from the current escapement goal, which was established in 2015. The datasets included in `EGprocess` represent this situation. 

## Spawner-Recruit Plot

EGPIT decided spawner-recruit (SR) relationships should depicted the SR pairs used to estimate the relationship (with color coding to indicate SR pairs included in the original analysis and added to the updated analysis), the estimated SR relationship, $S_{MSY}$, and the current goal range[^1]. 

[^1]: Note on the multiplier argument. The multiplier argument is required for Shiny app posteriors retrieved prior to May 2026 in which a scaled version of $\beta$ is included in the posterior distribution provided by the app. After May 2026, $\beta$ is provided on the natural scale and setting the multiplier argument to 1, which is the default, is appropriate.

```{r SRplot_nochange, fig.width = 6, fig.height= 6}
brood_Igushik <- get_brood(run_data = data_Igushik)
post_Igushik <- c(post_Igushik_byr63_05, post_Igushik_byr63_15)

plot_SR(posterior_list = post_Igushik, 
        brood_data = brood_Igushik, 
        goal_data = goal_Igushik, 
        title = "Igushik River Sockeye Salmon", 
        multiplier = 1e-5)
```

## Expected Yield Plot
EGPIT decided expected yield plots should depicted the empirical data used to estimate the relationship (with color coding to indicate SR pairs added to the dataset since the last time the escapement goal was changed), the estimate of expected yield, and the current goal range. 

```{r ey_nochange, fig.width = 6, fig.height = 5}
plot_ey(posterior_list = post_Igushik, 
        brood_data = brood_Igushik, 
        goal_data = goal_Igushik, 
        title = "Igushik River Sockeye Salmon", 
        multiplier = 1e-5)
```

## Optimal Yield Profile

EGPIT decided escapement goals intended to maximize sustained yield should use a universal standard target yield of greater than 90% of maximum sustained yield (MSY). When producing OYP plots this means a single line, overlain with the current escapement goal is needed to describe the probability of achieving that target within the escapement goal range. By displaying the OYP generated from both the original and updated analyses ADF&G staff can easily compare how the empirical data added since the last escapement goal change has impacted the relationship between MSY and the goal. In this case, 10 additional brood years resulted in a slight rightward shift of the OYP and minimal changes to the probability of achieving at least 90% of MSY at both the upper and lower bound of the existing goal. 

```{r OYP_nochange, fig.width = 6, fig.height = 6}
profile_Igushik <- get_profile(post_Igushik, multiplier = 1e-5)

plot_profile(profile_Igushik,
             goal_Igushik,
             "Igushik River Sockeye Salmon")
```

## Spawner-Recruit Parameter Table

EGPIT determined the following parameters should be presented after each escapement goal update. The table shows medians for each statistic along with the coefficient of variation[^2] and the 95% credibility interval.  

[^2]: The geometric coefficient of variation is $cv = \sqrt{e^{\left(\frac{log(q95)-log(q5)}{2*1.65}\right)^2} - 1}$ where q95 and q5 represent the 95th and 5th percentiles from the posterior distribution of the parameter, respectively.

```{r SRtable_nochange}
table_SR(post_Igushik[2], 
         title = "Igushik River Sockeye Salmon", 
         multiplier = 1e-6)
```

# Finding: New Escapement Goal Range

During the escapement goal review process for Igushik River Sockeye salmon the escapement goal review committee elected not to change the escapement goal. For illustrative purposes we will imagine they had recommended an escapement goal change and examine the changes to `EGprocess` outputs in that scenario. 

Imagine the committee had disregarded the stability of the relationship between the escapement goal bounds and the OYP probabilities displayed in Figure 5 and decided the goal was too aggressive relative to a target of 90% of MSY. This imaginary committee may have decided increases to both the lower and upper bounds of the existing escapement goal were appropriate. This decision could have been communicated using `EGprocess` outputs by adding an additional row to the `goal_data` file containing the new escapement goal finding and the year the new finding would go into effect.

In addition to the modified `goal_data` argument, the argument `new_finding` is set equal to TRUE for the following plots.  In what follows we will also change the figure titles to make clear that this is an imaginary example.

```{r}
goal_imaginary <-
  goal_Igushik %>%
  bind_rows(
    data.frame(yr = 2024,
               lb = 175000,
               ub = 500000)
  )
goal_imaginary
```

## Spawner-Recruit and Expected Yield Plots

When a new escapement goal finding is made the spawner-recruit plot is identical except the new finding is displayed on the plot and highlighted using color coding. Color coding is extremely helpful here as the new finding can be difficult to detect unless the reader remembers the current and new escapement goal bounds very accurately. The expected yield plot in this scenario (not shown) would have the same changes with the same modifications to the arguments of `plot_ey`.

```{r SRplot_change, fig.width = 6, fig.height= 6}
plot_SR(posterior_list = post_Igushik, 
        brood_data = brood_Igushik, 
        goal_data = goal_imaginary, 
        title = "Imaginary River Sockeye Salmon",
        new_finding = TRUE,
        multiplier = 1e-5)
```

## Optimal Yield Profile

When a new escapement goal finding is made the new finding is overlain on the optimal yield profile for the updated analysis while the current escapement goal range is overlain on the optimal yield profile from the original analysis. This configuration allows staff to compare the degree of similarity between the new and current escapement goals, relative to the analysis responsible for each finding, with respect to maximizing MSY. In this case, the new escapement goal finding offers improved symmetry between the upper on lower escapement goal bounds and roughly similar probability of achieving at least 90% of MSY at the lower bound but decreased probability of achieving at least 90% of MSY at the upper bound. 

```{r OYP_change, fig.width = 6, fig.height = 6}
profile_Igushik <- get_profile(post_Igushik, multiplier = 1e-5)

plot_profile(profile_Igushik,
             goal_imaginary,
             "Imaginary River Sockeye Salmon",
             new_finding = TRUE)
```

## Historical Escapement Plot

Note that `plot_escapement` will reflect new escapement goal findings as well if provided with `goal_data` where the last escapement goal change is later than the last escapement provided in `brood_data`.

```{r historicS_change, fig.width = 6, fig.height= 5}
plot_escapement(brood_data = brood_Igushik, 
                goal_data = goal_imaginary, 
                title = "Imaginary River Sockeye Salmon")
```

```{r, include=FALSE}
#Run this in the console when changes are made to update the html stored in docs.
#rmarkdown::render("./vignettes/EGprocess-Standard_process.Rmd", output_dir = "docs")
```
